\documentclass[12pt]{book}              % Book class in 11 points
\parindent0pt  \parskip10pt             % make block paragraphs
\usepackage{hyperref}
\usepackage{wasysym} % Needed for cents symbol
\usepackage{float}
%\usepackage{footnote}
%\makesavenoteenv{tabular}
%\makesavenoteenv{table}
%\raggedright                            % do not right justify

%\newcommand{\supth}{$^{\mathrm{th}}$}

\title{\textsc{Family Portrait}\\ The Memoirs of James Alfred Morris}    % Supply information
\author{J. A. Morris}              %   for the title page.
\date{New York, 1962}                           %   Use current date. 

% Note that book class by default is formatted to be printed back-to-back.
\begin{document}                        % End of preamble, start of text.
\frontmatter                            % only in book class (roman page #s)

\section*{Note from the editor}
This is an lightly edited reproduction of the original typed manuscripts of the memoir of J.~A.~Morris, with hand-written annotations and corrections. It was transcribed by Brett M.~Morris, great-grandson of the author, in 2014. 

The annotations in the margins have faded in places and are illegible. When large bits of text are missing that cause breaks in the narrative, they will be marked with the word: [Illegible]. %Missing pages will be marked similarly.

In an attempt to remain true to the manuscripts, any descriptions or definitions added by the editor to clarify some of the more dated language are written in footnotes. 



\maketitle                              % Print title page.
\tableofcontents                        % Print table of contents
\mainmatter                             % only in book class (arabic page #s)
%\part{A Part Heading}                   % Print a "part" heading
\chapter*{Forward}                % Print a "chapter" heading
%Most of this example applies to \texttt{article} and \texttt{book} classes
%as well as to \texttt{report} class. In \texttt{article} class, however,
%the default position for the title information is at the top of
%the first text page rather than on a separate page. Also, it is
%not usual to request a table of contents with \texttt{article} class.
% 
%\section{A Subheading}                  % Print a "section" heading
%The following sectioning commands are available:
%\begin{quote}                           % The following text will be
% part \\                                %    set off and indented.
% chapter \\                             % \\ forces a new line
% section \\ 
% subsection \\ 
% subsubsection \\ 
% paragraph \\ 
% subparagraph 
%\end{quote}                             % End of indented text
%But note that---unlike the \texttt{book} and \texttt{report} classes---the
%\texttt{article} class does not have a ``chapter" command.

I have always had the urge to write something of longtime interest to others. During a long business career I did a lot of writing -- studies, reports, suggestions -- of the current interest only. They were communications that seemed very important at the moment. But I daresay very few could be found at this writing (January, 1962). If available, they would have little meaning now. Perhaps a suggestion on business policy is still timely and hauled out of the back of an office file for reference. Of, maybe, a letter to family or friends contained a helpful comment that warranted storing in that packet of memories in the attic trunk. However, there are few communications in the average person's life that stand the test of time.

This was emphasized very strongly in my attempt to reconstruct the family genealogy. The record goes back for more than 300 years. But it is a record of births and deaths with a bare sprinkling of facts on which to evaluate the character, life and works of the individuals. Perhaps the urge to set down some words that would be of enduring value as a family portrait sketch for the younger offshoots is due to the fact that I, like others of my generation, will also be a mere statistic before many years. This urge to preserve a background of the family is much stronger than my lifetime reluctance to look back. 

[Illegible]

So, here is the beginning of a story that might, I hope, give life to some of the statistics on the recent generations of my family. Because it must necessarily revolve around the one I know most about, it is written in the first person. 

\chapter{Early Childhood: ``The Story Begins''}

For me, the story begins with a statistic -- April 6, 1893 -- in a two-story and basement frame house on Buffalo Avenue, Brooklyn. Around the corner at 844 Herkimer Street in the backyard garden of a more pretentious brick house, my father's young cousin, Caroline Halliday, was playing and keeping a watchful eye on the upper rear windows of the Buffalo Avenue House. She was watching for a signal that her mother, Grandma Morris' sister, was needed. The signal, a sheet hung out a window, came and Aunt Addie rushed to the Buffalo Avenue house and helped with my birth. Eighteen months earlier my parents had welcomed their first child -- Chester DeVere -- my big brother to the world. Four years later in 1897 my sister, Vada was born in the Buffalo Avenue house.

My father, Isaac J. Morris, was 25 years of age when I was born and my mother, Birdella LaBonte Morris, 23. Whether they wanted me so soon after the first child is a moot question. But I came along anyway and Mom told me in later years that I was a cry baby, a thumb sucker and an apron-string hanger-on until I was old enough to go out and rough-it-up with my playmates. I do know that I was strongly attached to her. 

I can still recall the misery experienced when she left me for the first time to visit her parents in Albany. We lived in Morris Park then and the memory of my walk with her to the railroad station is vivid. I was crying and she would pause to dry my eyes and console me. To this day, the sound of a train whistle in the night brings back the acute feeling of loneliness I felt that first night she was away. 

I have no recollection of my babyhood in the Buffalo Avenue house. All of my memories of the neighborhood stem from childhood visits to Herkimer Street with my parents after we had moved to Morris Park. Aunt Addie was a widow with three children -- Isaac, Carrie and Walter. Grandma Morris (Ann), also a widow, lived next door with her brother, David Swayze. At that time the neighborhood was upper middle class. To us it seemed like millionaire's row for we were quite poor. The street was cobble-stoned and I can still hear the click-click of horse shoes and the lurching screech of iron-rimmed wagon wheels on the stones. 

The gardens in the rear yards stretched back for maybe 150 feet and we enjoyed playing in them. There were a couple of peach trees with low-hung branches that tempted us to snatch at the big, luscious fruit when Aunt Addie could not see us. 

The Haliday house was a lively one. The two boys were always playing pranks on their mother and sister, and on Chester and me when we visited them. Cousin Ike was the older and more serious of the two and highly opinionated -- a fact that made him rather forbidding in later life. Walter and Carrie were really fun-loving and laughter was the order of the day -- even more so when Arther Blinn, Pop's nephew, visited the Halidays. I think Arthur's mother, Caroline Morris Blinn, died during his birth. It was a home we loved to visit. But it was not without its tragedy. 

Aunt Addie had the problem of raising three children. To help bolster a limited income, she took in boarders. One of them, a handsome real estate broker, courted and married her in 1901. A short time later he disappeared and was not heard of again. Although she never said so, we believe she expected he would return and for that reason has remained in the Herkimer Street house. At this writing, she is well into her eighties and still there in a much run-down neighborhood -- a sad contrast to its stately refinement of her childhood. 

Despite little formal education, Cousin Ike was successful in business. He was associated with an estate and at one time as part of his responsibilities managed the \href{http://en.wikipedia.org/wiki/Hotel_Theresa}{Hotel Theresa} in Harlem. That section of Harlem, now entirely negro, was a very fine neighborhood. It was really a big treat to dine with Cousin Ike and his wife Birdie. Mother would polish us up, put on her very best clothes and admonish us on our table manners. Cousin Ike later became night manager of the New York Stock Exchange Clearing House, a job he kept until his retirement. 

Grandma Morris, Uncle Dave and Aunt Addie have long since passed away. More recently, Cousins Arthur, Ike and Walter have died. Walter's son, Walter J., is living in Rockville Centre. 

As mentioned earlier, I have no recollection of my first home in Brooklyn. We moved to Morris Park (\href{http://en.wikipedia.org/wiki/Richmond_Hill,_Queens}{Richmond Hill}, south of the railroad) when I was about four years old. Pop was a machinist in the Long Island Railroad shops in that village. The new home was a far-from-pretentious flat in a three-story six-family house on the northwest corner of Briggs Ave (now 117\textsuperscript{th} Street) and Chichester Avenue (now 95\textsuperscript{th} Avenue). There was only one house directly opposite on the avenue. In the rear of our flat was a fenced-in yard and beyond that open fields almost to Atlantic Avenue. This is where we played ball when we were older. 

Morris Park was really a country village in those days. Five blocks south, Liberty Avenue was ``the end of the world.'' From there we walked through the woods (now Glen Morris) to the old water holes in the swamps of the Aqueduct. And we did not need bathing trunks for our swimming, the high grasses and cat-o-nine tails\footnote{a.k.a \href{http://en.wikipedia.org/wiki/Typha_latifolia}{\it Typha latifolia}, a wetland weed} hid us from view even if there was any one within hailing distance. 

Two events are the earliest recollection of my childhood. Both occured at the first home in Morris Park. I was just a little toddler playing in the yard while Mom hung the wash. An Irish woman leaning out an upper window shouted that my pants were falling open (actual words censored). If a baby can be embarrassed, I was and so was Mom. It was peculiar that an incident like that can be remembered while other more important ones cannot. The other event was at a Thanksgiving or Christmas party in our flat. The Halidays and Grandma Morris were there and my little baby sister Vada was the center of attention. In some way, one of the glass dessert dishes was cracked and I swallowed a small chip of glass. Pandemonium broke loose with much shouting and wringing of hands probably brought about by my wailing. Finally some one thought of calling a doctor. He filled me full of crackers or bread, topped off with a large dose of castor oil. Evidently they found the glass. 

\section*{Pop}

Pop was an apprentice machinist in the railroad shops at Albany or Rensselaer when he courted Mom. [Illegible] I do know that he worked hard and had long hours. As I recall, the deep-throated, long-carrying shop whistle awakened the entire village at seven each morning, blew at 12 noon for a half hour lunch period and at 6 p.m. for closing. That was a 10.5 hour day for the shop personnel. His pay check for a six day week was \$12 or \$15. Compared with present day hours and pay that certainly was unbelievable. 

It was a monotonous treadmill of work with little time left for anything else and with no future. Pop was handicapped by lack of formal education. He had not completed grammar school but was an excellent mechanic and of an inventive turn of mind. He sought to lift himself through inventions and by going into business for himself. He was not successful in either direction. 

At one time he felt there was a market for a convenient hand washing compound. From his daily experience he knew the difficulty of washing grease and grime from his hands at the end of a day's work. His idea was a compressed washing powder about the diameter of a quarter and a quarter inch thick, packed in rolls like the old candy \href{http://en.wikipedia.org/wiki/Necco_Wafers}{Necco Wavers}. Held in the palms of the hands under the water tap it would dissolve and with a little rubbing remove the grease. He advertised for shop workers with little success. [Illegible] Later the big soap companies introduced hand washing compounds in pase form which met with the success that Pop had so much sought. 

Another of his many inventions was a collapsible wooden crate for packing onions and other vegetables -- that could be returned to the farmer or produce dealer and reused. He patented the idea, incorporated a company under the name {\it American Crate Company} and sold stock to his friends. Then the promoter sold Pop on teh idea that he could sell the crate to the onion growers of Texas and departed with the remaining cash. Years later he sent shares of stock in an oil exploration company to Pop and to the stockholders of the Crate Company. They turned out to be as worthless as the crate company stock and ended another dream for Pop. 

We can return to our early days in Morris Park with Pop's venture in the retail grocery business. Three doors down the block from our flat was a little frame house, the front room of which had been converted into a store. Pop left the railroad, moved the family into the little house and opened a grocery store advertised as ``The Little Store with the Little Prices.'' One thing I recall about the house was the grape arbor forming a leafy corridor to a two-holer outhouse. Perhaps the corridor was not too long, but one time I didn't quite make its entire length much to the consternation of Mom and my own embarrassment. By way of extenuation it may be said that I was only four or five years old. 

A very short time later, Pop arranged to have a new store built on the lot immediately adjoining the little store. The structure was built on the entire width of the lot. It had a built-in driveway to a stable in the rear of the house. The store and rear storeroom took up the remainder of the first level. From the storeroom a staircase led to a very nice apartment above. Pop continued in business here for a few years until his leniency on credit brought the business toppling down.

An old oder book (I don't know if it was for delivery from the little store or from the new one) discloses some rather startling prices when compared with those today. Here are a few examples: 

\begin{table}[H]
\centering
\begin{tabular}{l r}
3.5 pounds of sugar & 17\cent \\
1 pound of coffee & 36\cent \\
1 loaf bread & 5\cent \\
1 can condensed milk & 10\cent \\
1 quart milk & 5\cent \\
1 gallon Kerosene oil & 13\cent \\
1 lamp wick & 2\cent \\
1 quart white onions & 12\cent \\
1 peck potatoes & 18\cent \\
1 bushel coal & 17\cent \\
\end{tabular}
\end{table}
%\footnotetext{1 peck of potatoes= 15 pounds}
%\footnotetext{1 bushel of coal = 80 pounds}

\chapter{Memory begins to crystallize}

It is from this place that my memory begins to crystallize. Names of playmates, the rough and tumble fights of childhood, the short cut across the fields to the primary school on Elm Street, and the many other things that made their impressions on a developing mind. I would like to mention a few incidents not only for the nostalgic interest they hold for me but as a sort of factual background for appraisal of the effect on our character. 

First, let me say that my sister Vada was born in this house in 1897. I don't remember very much about her entrance, only the few times Mom asked me to push the baby carriage. I do remember the admiration she drew from the relatives when they visited Morris Park. She was a beautiful baby and grew into a beautiful girl and woman.

Because Pop's name was Issac, two of the neighborhood boys delighted in calling me ``Ikey.'' One of the boys, George Washington, was colored and the other, Tommy Givins, Irish. They were a little older and I was dreadfully afraid of them for a long time. Finally the taunts of Tommy hurt so much that I forgot I was afraid and we fought it out in the fields coming from school. At the end we were both sobbing and I may have won by a very small margin. Later I managed to fight it out with George also. With those two childhood fights came confidence in my physical ability to give and take -- a confidence that saved me from many other boyhood fights. I had learned that a bully thrives on the fears of his victim. Take away the power to frighten and the bully is deflated. 

I learned, however, through a different kind of incident that there are other kinds of fears. For some reason or other my teacher in old 53, I think it was the first grade, locked me in a closet. I was frightened to the point of hysterics. When she opened the door, I was on the floor with my nose to the thin strip of light at the bottom of the door. For many years after, I was afraid of the dark and to this day fear closed places. 

Our new home was lighted with gas continuously by ``pay-as-you-go'' meters. One Sunday evening the lights went out and Pop asked me to go to the cellar and drop a quarter in the meter. To do so I had to go downstairs, through the dark store room and on down to the even darker cellar. I was searching for the slot in the meter when something scrambled away. I ran upstairs screaming and Pop went down and discovered a cat in the cellar. That was another experience that made me afraid of the dark.

Mom was deeply religious and her training was reflected in our early formation of character. All of her social contracts were made at the First Methodist Episcopal Church of Morris Park and her children were drawn into the circle. On Sundays we attended the morning services with her. We could be found at Sunday school in the afternoon and after supper at the evening service with Mom. I can still recall sitting beside Mom and singing such hymns as {\it Jesus Lover of My Soul}, {\it Lead Kindnly Light} and {\it Rock of Ages}. When we were a little older we joined the Epworth League and attended those meetings just before the evening service. Sunday was indeed a crowded day. 

Wednesday was also a special day for Mom. On that evening she went to prayer meeting with a close friend and neighbor, Mrs. Bradford Wicks. Pop rarely attended any of the church services. 

Mom did not believe in any frivolity on Sundays. She would not let us play ``catch'' or any games and it would have taken a real emergency to allow us to ride a trolley car on that day. Sunday was a day we wore our blue serge suits and didn't dare get them soiled. When they became too shiny for Sunday and holiday we used them for school. 

The Methodists were really strict in those days. Each service was an evangelistic meeting in itself with much talk of hell and brimstone. Dancing was frowned upon and drink was a curse. Mom would get real excited when a Catholic neighbor would pass the house with a tin pail on the way to a nearby saloon for a pint of beer. This was called ``rushing the growler'' and the tin pails were large enough to hold far more than a pint. She could not understand either why the Catholics would permit ball playing on Sundays or hold picnics or other festive affairs. 

With no movies, radio, television or automobiles, the church -- no matter what the denomination -- was the social center of the community. Apart from religious services, in the summer there were strawberry festivals on the lawns of the church or private homes, trolley rides to Coney Island and picnics. In the winter there were straw rides in horse drawn sleighs or wagons, or house parties of the various societies.For the children there were two big events. One was the Anniversary Day Parade in early June with ice cream and cake awaiting them in the church basement at the end of the march. With bands playing such tunes as {\it Onward Christian Soldiers} and with the children decked out in their best clothes carrying banners and American flags, it really was a gala event. 

The other big event for children was the Christmas party a few days after Christmas. Following the entertainment, Santa Claus handed boxes of candy and oranges to the children. Sometimes there would be presents from the teachers. One of the then popular G.A.~Henty books, dated December 25, 1902, in my library was a present from my Sunday School teacher, Skidmore Pettit, Jr. 

Kids in that far-away age were no different that kids today, particularly when food was concerned. When refreshments were served on Anniversary Day, there was keen competition on the number of dishes of ice cream we could get away with. The winner was usually the one who could sneak back in line the most times or who would have some friend hand out through a basement window dishes of the stuff purloined from the tables behind the line. 

The same Christmas I received the Henty book, I was presented with a small volume New Testament inscribed ``Presented to Alfred Morris for faithful attendance at the Sunday morning church service by the Sunday School of the First Methodist Episcopal Church of Morris Park L.I.. W.C.~Van Horn, Superintendent. Lincoln H.~Casweel, Pastor. Christmas MCMII.'' 

In the eyes of a nine year old boy, Dr.~Caswell was an old man, but he was probably 35 or 40 years of age at that time. Like his first name, he was a kindly gentleman with a ready smile and made all feel at home. The Sunday School Superintendent, Mr.~Van Horn, also led the singing. His leather-lunged voice is etched in my memory. It literally rattled the rafters and was infectious in its demands on our own vocal chords. He certainly was a top-notch song leader and a fine Christian. 

The preached succeeding Dr.~Caswell as a Dr.~Chadwick, a large, raw-boned elderly man with thick grey hair and bushy sideburns. He hammered home the gospel in a resounding voice. More than any other I have heard, he typified the old-time methodist preacher who believed and preached eternal damnation for the sinner. There were other ministers that followed but non like him. One, in particular, when I was a young man really shook my faith. He deserted his wife and ran away with the voluptuous vocalist of the choir -- the wife of the master of my masonic lodge. It was some time before I returned to the church after that incident. 

Aside from my mother and father, the person whose teachings did most in moulding our moral and religious concepts was our Sunday School teacher for many years, Skidmore Pettit Jr. He started a class of eight to ten year old boys and continued with the group until we were well into manhood. Chester and I were members from the beginning. By words and deeds he instilled in us the urge to be manly, clean and honest. To better hold us together he formed a social club, the C.I.O.C. (Christ Is Our Captain) which alternated in meeting at the homes of the boys. We wore pins inscribed with the letters of the club and were proud to be members of the class and the club. At one time I was president of the club. But this activity, like so many others in life, was not without its tragedy. 

Unfortunately, Mr.~Pettit's teachings evidently did not penetrate the understanding of two of the boys. One became a drunkard and another was indicted for theft. One of his sons, probably the best of the class, was killed in action in the First World War. Most of the boys became respected citizens and many successful business men. 

\begin{center}
------------------------------------------------------
\end{center}
In penning a story of this type, one's thoughts come more quickly than they can be spelled out and account somewhat for apparent digressions. But the object is to present a background to help bring statistics to life. And in order to do so we have to tell something about my youth, business life, courtship and marriage to Mabel, my son, Jim, his wife, Vandy, and our two grandchildren, Scott \& Karen. Then there are my brothers and sister, etc -- and there is more to tell about the early days of the family in Morris Park. 
\begin{center}
------------------------------------------------------
\end{center}

% Timeline: April 1903

Pop continued in the grocery business until the latter part of 1903 and left it with a lot of money owing him and a lot of debts. He returned to work in the railrad shops and continued there until his retirement in 1921. We moved to a duplex house on Sherman Street. It was very much run-down and had no gas or central heating system. We used kerosene lamps and the kitchen range to provide most of the heat. Little portable cylinder-shaped oil stoves took the chill off the upstairs bed rooms but always with the danger of explosion or being tipped over. We were all very unhappy in the old house.

This place was less than a block from the railroad tracks running alongside the railroad shops. In those days the railroad was not electrified and the firing of the steam boilers on the moving locomotives would scatter pieces of coal on the tracks. I can recall walking the tracks with Chester hunting for such coal and half-burnt cinders. We would put them in a bag and drag them home to help feed the kitchen range. We were not asked to do this but we felt it was a contribution that might help Mom put a few extra pennies in the old cracker jar to buy needed shoes and clothing for the family. 

It was experiences such as that that made a deep impression on us and developed the will to strive for success. A will that was further stimulated by avid reading of \href{http://en.wikipedia.org/wiki/Horatio_Alger,_Jr.}{Horatio Alger, Jr.} stories. They were the ``rags to riches and poor but honest'' type of fiction which in my opinion had much to do with building character in the youth of that day. Although considered trite and laughed at today, I believe that, next to our religious training and close family ties, the writings of Alger had a lasting influence on the moral concepts of the ambitions of all who read them.

After a short time in the Sherman Street house, the folks rented a beautiful little one family home around the corner on Chichester Avenue (now 95\textsuperscript{th} Avenue).  It was only a short distance to Public School 57, which we attended. We were very happy in that house. Then we moved to a large house on the corner of Beach Street and Belmont Avenue. This was only a block or so from Adikes farm and I can recall going there after the cows had returned from pasture for pails of fresh milk. Nearby lived a little freckle-faced playmate, \href{http://en.wikipedia.org/wiki/Percy_Crosby}{Percy Crosby}, who would later become a famous cartoonist, the originator of the syndicated comic strip ``Skippy.'' 

% Timeline: 1905

By 1905 or 1906, Pop's finances were evidently in better shape than they had been for many years. He contracted with the local builder (Jeffreys) to build a two family house at 2 Briggs Avenue (now 911 117\textsuperscript{th} Street) just up the avenue from our first home in Morris Park and two doors north of the rear property of the Catholic Church. We occupied the second floor and Mom made a very comfortable home of it. For us children at least, it became the ``old homestead.'' I lived there until my marriage in 1920.

We were all very happy in the new home at 2 Briggs Ave. It was here that we spent many enjoyable years through our teens and into adulthood. Our brother, Kenneth, was born in this house. On the morning of September 19, 1908, Pop asked me to play in the back yard. A short time later, he opened a rear window and shouted ``it's a boy!'' So with the ``kid'' brother added to the family, Mom and Pop were proud parents of one girl and three boys. 

Looking back, Mom must have had a difficult time during her pregnancy. It had been seventeen years since Chester was born. Childlike, we didn't understand her spells of depression and her tears at the slightest provocation. Later, when we were raising our own families we could appreciate the trying period she went through. As later events proved, Kenneth's entry into the family was the source of much happiness and contentment for Mom and Pop. This was particularly true in their declining years with the three other children married and away from home.

Many, many pleasurable childhood memories are associated with the Briggs Avenue home. Playing marbles on a warm spring day under the big elm trees on the Atlantic Avenue side of the open field adjoining our house, jumping on the rear step of a passing horse-drawn ice wagon for a few chips of ice to suck on, climbing and hiding in the trees to puzzle passersby with water from our water pistols, playing ball in the open fields and on weekends sitting on the grandstand of the ball field just across the tracks watching our local heroes ``The Tigers'' play. In the winters there was the usual belly-wopping\footnote{sleigh riding on one's belly, according to {\it British-American Dictionary} (1996) by Catherine M.~McCormick} and playing in the snow. Mom would worry about our always-wet feet and when we returned she made us sit in front of teh kitchen range with our feet on the edge of the open oven door. Sometimes we would take off our shoes and put them in the open oven for a quick dry. If we left them there too long the toes would curl up and make it pretty uncomfortable when we put them on again. The smell of wet leather is still reminiscent of those early winter days. 

Then there were the chores -- some not so pleasurable. The new home was heated with a coal burning furnace in the cellar which made a lot of ashes. On Saturday mornings it was our job to sift the ashes for pieces of half-burnt coal that had dropped through the grate. These were used to bank the fire at night. The sifting kicked up a lot of dust and we really needed that weekly Saturday night bath.

I can recall also on each of Pop's pay days taking the mortgage payment book with some money in an envelope to Mr.~Jeffrey's house across from the Methodist Church. When Pop had to work overtime, Mom would prepare a light supper in one of those old workman's dinner pails. I would take it to the shops passing other workmen on their way home. Pop worked on a huge lathe grinding worn engine and car wheels. At times, Pop would be in the midst of a job and could take time only to shout a hearty ``hello!'' I liked that errand particularly when he had the time to chat and tell me about the machines. 

Down the avenue in the old flat building, our first home in Morris Park, the ground floor corner apartment had been converted into a grocery store owned by Henry Voige, a German. Across the avenue on the opposite corner was Dusenberg's candy store where for a penny we could buy a bag full of candy. I can't say it was very wholesome, but what it lacked in quality it made up in quantity. When I was about 12 or 13 years of age I started working on Saturdays at Voige's store.

The Saturday workday was from eight in the morning until eight or nine at night for the magnificent sum of 50$\cent$. And it was hard work delivering big boxes of groceries and bushels of coal also. Many people with only kitchen ranges for heat would buy coal by the bushel in those days. The coal would be dumped in a box behind the range. At the end of the day I would be very tired. I remember one particularly heavy day when I walked home sobbing from exhaustion. I had the feeling that the job was worth more than 50$\cent$. I put the problem to Mr.~Voice, but when he told me that he could get other boys to work for that money, and I didn't have to work for him if I didn't want to, I told him I would be glad to continue as I liked the job. Later, I did get a raise. 

During school vacation the work was not quite so hard but in some respects not to my liking. In the mornings, I would clean the stable and hitch the horse. To a person who has never cleaned a stable by sifting the dirty straw bedding of fresh manure and then raking the straw into a heap for the night's spreading, the experience would be revealing. All that is needed is a pitch fork, a rake and a long-handled shovel, as well as a clothes pin for the nose. The long-handled shovel is needed to scrape up the residue from the sifting, fill it with the stuff and carry it to the manure heap in the yard. One must be sure when leaving the stable to scrape the bottom of the shoes with the sharp end of the shovel. Otherwise the store and the home might capture some of the delightful aroma of the stable. 

Usually I did not mind driving the horse and making deliveries. It was fun and I was a bit proud at being a wage earner. But, when on a Sunday School Anniversary Day, I had to wait on a corner in sight of the paraders and then had to cross the line of march, my ego was really depleted. For these little girls in their holiday frocks and my pals in their blue knickers and white shirts to see me on a wagon in my old duds was the height of embarrassment. 

Friday evenings at the store we would work overtime (but no overtime pay) weighing sugar into 3.5 pound quantities and pouring it into bags until the whole counter was filled. Then we would fold the tops of the bags and pile them for Saturday rush. There were few package goods in those days. Most of the goods were in bulk and weighed on order. Wooden boxes of prunes and apricots were open on one end of the counter. In front of the counter toward the back of the store there was usually an open barrel of sugar, one of mackerel in brine and another of assorted crackers or those ball-like spiced cookies. The open barrel of crackers was real handy when the male neighbors would pause in their shopping for a bit of gossip. Then we would hear some timely cracker-barrel philosophy which to me was the wisdom of the sages. 

\section*{Mischief}

Mention of barrels reminds me of election night celebrations. As most bulk goods were packed in barrels, used ones were very plentiful. A week or two before election the boys would start ``collecting'' barrels from the backyards of houses where they were used for rubbish and other building materials, from the yards and other outdoor storage spaces of shops, and from any other places where the owners had perhaps forgotten that it was near election day. At times we would have a collection of 20 or 30 barrels for the big event.

After dark on election night they were taken to some open lot and pyramided as high as possible and ignited. The resulting blaze was our contribution to the celebration of election victors, regardless of party. At our celebrations, one could turn in any direction and see the sky lighted with other such fires of patriotism. Peculiarly, the election celebration of ``borrowing'' and burning of barrels was apparently condoned by our parents for I can remember no time when parental action was taken to stop it. It is true however that the owners of the barrels did not like losing them.

Halloween was another day when the boys stepped a little out of line. For reasons which I cannot understand, the prevailing practice among the kids was to fill the feet of their old long cotton stockings with flour and sock anyone within range of the swinging ``fun-maker.'' As soon as we left school, we would turn our coats inside-out, get our flour filled stockings and star swinging. If one's clothing was not completely white at the end of the day, we were not having fun. Some of the tough kids would be somewhat vicious by using lime or dirt in the stockings and take delight in the punishing blows they could administer. On Halloween evenings everything portable had to be securely anchored. Otherwise the boys would steal gates off fences or take anything moveable and leave them in some neighbor's yard perhaps a quarter mile away. I was unfortunate one night when our crowd started to pull a hearse across the fields and onto a road. I was the one caught. The masked ragamuffin outfits and begging at doors was reserved for Thanksgiving day. That was the day the younger children had their fun.

Such pranks are worth telling if only to show that, even if practice differs, the kids of yesterday are no different than the kids of today. However juvenile delinquency as such is infrequent or at least the expression was seldom used. We were either good boys or bad boys and their seemed to be more good ones than bad ones. In my opinion, this less pronounced delinquency problem was due to more closely-knit family life and more strict parental discipline, closer church ties, tougher police (they used to carry night sticks), and plenty of open spaces for play.

In our area there was ample room for baseball, tennis, and all outdoor sports. There were nearby hills for coasting and tobogganing and ponds for fishing and skating. The growth of population in urban and suburban communities eliminated our playing room and gave rise to gang growth. These groups of boys drawn together by boredom seek an outlet through bullying and crime -- a condition fostered and strengthened in part by the horror of three wars, the portrayal of crime through movies and television, and the decline of the church as a social center of community. 

In our late teens, for example, a group of about eight of us received permission to build two tennis courts on Belmont Avenue. We scraped and rolled the courts, built a wire enclosure, erected a judges stand and player benches and then spent our idle time in strenuous play. I was elected the first president of teh Belmont Tennis Club which was in existence for about four years. 

There are many memories of early childhood the recital of which would leave little time and space for the interesting parts of our later family portrait.

\chapter{A New Era}
% Timeline: April 1908

The year 1908 was an eventful one. It opened a new era. As I mentioned before, Kenneth was born in that year. Chester had left school and was working in a textile firm in New York. Vada was in grammar school and I was graduated from grammar school. A sad event was the death of Grandpa LaBonte in Albany where he lived for many years. I had not seen him very much and really knew little about him. My impression is a stocky built man with a large white mustache. He was a conductor for many years on the Boston \& Albany Railroad. Chester and I inherited his prized Columbia Bicycle. Other than makeshifts, it was the first real bicycle we owned. 

In September 1908, I enrolled in Commercial High School on Albany Avenue, Brooklyn (now \href{http://www.brooklynvisualheritage.org/alexander-hamilton-high-school}{Alexander Hamilton High School}). The decision to attend the Brooklyn school was my own. Richmond Hill High offered only a four year academic course. I was raring to get into business and the three year commercial course offered by Commercial High appealed to me. It seemed to offer those subjects that would help me get started more quickly. The course included stenography, typing, bookkeeping, business arithmetic and courses of that type. I could not have taken advantage of this opportunity had it not been for the fact that as a student son of a railroad employee I was entitled to a pass on the Long Island Rail Road. Getting of at the Nostrant Avenue Station, I could walk the mile or so to the school, so there was no additional expense involved. 

But it took me 3.5 years to complete the course. I simply could not master the foreign language subject, German, which I had chosen to take. After I had flunked out in German, I took shop work and came through with average marks, graduating in February 1912. For a short time, I worked as a bus boy during the lunch period in the school cafeteria. During a good part of my high school days, I continued to work at the grocery store on Saturdays and during vacations. 

Those days at old Commercial were happy ones. Maybe I did worry some about my studies and do a lot of cramming to make the grade but those are not the things we remember most. It is true that we can recall that pride of accomplishment that came with some effort we made. However, like in all periods of life, our most vivid memories are concerned with the people we associated with the friends we made. 

Of all the personalities I met at high school, one stands out. He was Wad Smith (Edgar Wadsworth Smith), a schoolmate I met in 1910 when I was 17 and he was 16. Out of this meeting grew a friendship of fifty years. Wad was my closest friend and confidant. Years later we were to become business associates. Wad was a philosopher. He had an analytical mind, a retentive memory and the ability to express his thoughts clearly either orally or in writing. He was an avid reader and an artist of sorts. When we first met, he was Art Editor of Commericial's monthly magazine, The Ledger, and a contributor to its editorial pages. It was through his urging that I contributed a story entitled ``Grit'' which was published.

Wad lived with his parents and a younger sister, Dorothy, on the third floor of a walk-up apartment at 135 Rogers Avenue, Brooklyn. Our friendship ripened through discussions on subjects of mutual interest on our walks together to the railroad station or his home to continue the talks, we found that we had much in common. I had the opportunity to meet his friends from the Bedford Presbyterian Church and he mine from the Methodist Church. But more about that later.

One marked difference between the high school boys (and college boys, too) of that era and those of today was the choice of apparel. Casual dress then was the exception. A suit and stiff white shirt collar was the only acceptable attire. And the collar was detachable from the shirt with matching holes in the front and rear of the collar and shirt neckband for fastening with bone or brass collar buttons. With such an arrangement, one could put a clean collar on a soiled shirt and still look dressed up. Sometimes the very necessary buttons would drop when changing and if they could not be found and no replacements were at hand, we were simply out of luck. To save laundering, a popular collar for the young fellows was made of celluloid. This could be cleaned with a damp cloth either on or off the neck.

Some how or other this prevailing practice on dress made us seem more mature. When we reached the age for discarding knickers for long pants, we were men. At least that is the feeling such a shift gave us. Most of us would keep a sharp crease in those long pants. I can recall just before a date standing in my underwear before an ironing board pressing my only good pair of pants under a damp cloth. Later, when in business and could not get home in time to change, I would go into a tailor shop, take off my pants behind a screen and thrown them to the tailor for pressing. But that was only on rare occasions because pressing by a tailor cost money and money was a scarce commodity. 

Sunday was the day we really put on the dog. I could be seen after Sunday School parading on Jamaica Avenue or in Forest Park in our dark suits, high stiff collars and derby hats (straw hats in the summer, for a fellow wasn't dressed without a hat). Some of us would even wear spats and carry canes. Likewise the girls were out in all their finery -- long dresses and big picture hats and fancy parasols. Later we would congregate in Harsch's Ice Cream Parlor on Jamaica Avenue. If our flirtations had met with success we would treat our new girl friends to big 5$\cent$ ice cream sodas and then walk home with them. On the other hand if we were short of change we would wait until the girls had bought their own sodas before approaching them. Such an approach however was not always successful. 

The usual meeting place for the boys was at the home of the Aston's on Oak Street, the Morrow's on Jefferson Avenue or our home on Briggs Avenue. At the center of attraction at all places was a piano. Both Ed and Tom Aston played and wherever they were their playing would start the boys singing. Neither Bill nor George Morrow played, but their sister Ethel did. At our home Vada played but a difference of four or five years in age is a wide on for teenagers so she was not included in our impromptu sings. A good pal of mine, Art Wenige, played by ear practically any song requested, and he certainly could ``rag'' them. I can see and hear him yet, bounding up the front stairs of our home, making directly for the piano and shaking the room with his ragtime. He and I had a lot of fun composing songs -- he the music and I the lyrics. One, ``A Chapter of Beautiful Dreams'' we tried to have published, but could not break the barriers of Tin-Pan Alley. 

Like all boys, we had our ``bull--sessions'' with serious discussions on topics of the day but particularly on our plans for our future in business. My own ambition was to get into the advertising business. If we had recordings of those discussions they would be mighty interesting to hear today, although I venture to say they would differ little from those carried on currently by present-day youth.

The big mixed social events of our time were house parties with the boys and their girl friends again around a piano or dancing or playing various parlor games -- spin the plate, musical chairs, etc. Then sandwiches, ice cream, cake and coffee and the walk home with the girl friend of the evening. She lived considerable distance from my house, the return from her house to my house through the quiet streets was a lonely one. This was particularly so if the route took me near an isolated wooded area. Then anyone within hailing distance could hear me whistling to keep up my courage. 

It was a lot of fun, also, rehearsing and presenting minstrel shows at the old Arcanum Hall on Jamaica Avenue or the Temple Forum (built much later) on Johnson Ave. This yearly event was a big one which helped us to raise funds through our Knights of King Arthur society (KOKA) for the church.

To the teenager of today, all of this may seem tame. However, one must realize that we had to make our own fun, which we did and had a wonderful time doing so. Radio was not in general use, movies were just being introduced and television was still a long way in the future. The few automobiles were only for the wealthy and not dependable at that.

Outdoor sports were no different than today. We had our baseball team, the Mohawks with a field in Glen Morris, and as mentioned earlier, our tennis club and courts. It may seem odd but those two sports in which skill depended upon speed an keeping your eyes on the ball taught me a lesson which I have never forgotten and was a stimulating force in m business efforts. Years later when I was working desperately in a large corporation to overcome my lack of academic education in competition with college men, I would continually strive to keep my eyes on the ``ball''. In my dreams at night I would hear the impact of the bat on the ball and see the ball soaring into the outfield. I would awaken with the conviction that I could and must keep my eyes on the objectives of the job in hand and really it them hard. By doing that I would have no fear of competition. 

Occasionally Wad smith would bring his baseball team from Brooklyn to play our team. He and I also played a lot of tennis -- a game we played together well into middle life. His father had purchased a small summer bungalow in a new development in Allendale, Staten Island> The developers had built a community club house and a couple of tennis courts. Here we played literally from dawn to dusk on those weekends I was the guest of Wad.

For a couple of weeks one summer, Chester and I and six or seven of the Richmond Hill boys pitched a camp in the woods on the west shore of Lake Ronkonkoma. At that time, Ronkonkoma was very sparsely settled and the lake little used by tourists. It was very picturesque and if a dozen persons were seen on the beach near Hoyt's dock we would have considered it crowded. Another year I spent a vacation of a week or so at Comp Wopowog in Hampton, Connecticut. With Wad's crowd I also roughed it for a short time at a cabin on Cedar Lake near Denville, New Jersey. The cabin was loaned to us by the father of one of Wad's friends. 

We also had some good times at a large club house built on stilts at Goose Creek on the Rockaway Beach line of the Long Island Railroad. We would go there at the invitation of Harry Maass who was one of the Methodist crowd. Harry was an orphan and for very little pay while at school he would work on weekends cleaning up at the club. During the week, he was permitted to use the house. And he did in a big way by inviting our crowd to spend the night with him. The club had six or eight bedrooms all equipped so we were taken care of . I can recall the swims and sometime after midnight playing cards. The real thrill was diving from the tower at the end of the dock in the pitch blackness of night. 

All of the incidents and experiences mentioned so far occurred before our entry into World War I in 1917, some of them while I was in high school and some after I had started to work.

\chapter{Office Work}

My first experience in office work was with the \href{http://en.wikipedia.org/wiki/Frank_Munsey}{Frank A. Munsey Company} in \href{http://en.wikipedia.org/wiki/Flatiron_Building}{the Flatiron Building} in New York. It must have been in 1910 or 1911 during summer vacation that one of the boys suggested that a friend of his was going on two-week vacation and would appreciate my substituting for him as office boy. I did and was out of the office more than in because of the amount of messenger work I was assigned. As Munsey was one of the largest pulp magazine publishers at that time I received good bonuses in the form of copies of their magazines. I don't remember what pay I got. 

% Timeline: April 1912

Following graduation from high school in February 1912, I took my first job with the United States Title Guaranty Company in Jamaica. Frank DeBevoise, a fellow member of our church was employed by the company and asked me if I would be interested in coming with him. His older brother James was an assistant secretary of the company. I gladly accepted the job and was employed in the dual capacity of office boy and title searcher at a weekly salary of eight dollars. This was it. It was the beginning of a business career. I was happy. I could give mom five dollars a week toward the family income and still have three dollars a week for myself.

I did odd jobs about the office, helped with the photo-stating equipment and searched records in those very large canvas-backed alphabetically indexed books in the nearby county clerk's office. I enjoyed the work, my associates and the environment of a busy office. That is until I was tempted by a better offer by Harry Elsebough, our next door neighbor on Briggs Avenue. He was a bookkeeper and credit man for Walter W. DeBevoise (no relation to the DeBovoises in the Title Company), a candy manufacturer in 163 Carlton Avenue, Brooklyn. His offer of a job as assistant bookkeeper at ten dollars a week was just too tempting to resist. I accepted, left the Title Company on August 31 and on September 16, 1912 started my new job with DeBovoise.

On that morning I took a Long Island Railroad train to Flatbush Avenue and walked the mile or so to the factory and reported for work at 7:45 am. The factory was a three-story brick building about 200 feet by 150 feet. After the fairly modern office of the Title Company, the DeBevoise office was really antique. A better word might be old-fashioned. The office consisted of an eight foot long, two-sided bookkeeper's desk at which one could stand to work or sit on a high stool. A shelf elevated over the length of the desk was for ledgers or papers. At one side stood Harry Elsebough, coatless and with the green eye-shade on his forehead. The other side opposite Harry and behind me was a little corner sink with a mirror above it. 

Backed up against the front end of the ``desk'' was an old rolled-top desk for use of the city salesman. Back of Harry and near the front door was the combination typewriter and bookkeeping machine operated by a middle-aged woman. In one corner of the room was a walled-in cloak room and in an opposite corner nearest Harry and the typist, about the same size as the cloak room, was Mr.~DeBevoise's private office. This consisted of a large roll-top desk, a big leather-covered swivel chair, another cane-bottom chair, a wash bowl and a small wardrobe closet. The window ledge was used for filling odds and ends and for candy samples the boss was interested in. Add to this first impression the (at first) overpowering smell of chocolate and roasting peanuts and you have my introduction to the business in which I spent many years.

Here I was in my second unplanned job and one quite different than the first. Neither, however, were in the advertising field where I thought I belonged. I didn't get into that field for the simple reason that I didn't try to do so. Like so many boys of that era -- and of today, too -- I was opportunistic and took the first job offered because it provided an immediate income. Perhaps if my parents had the financial means to send me to college or I had the guts to make that opportunity myself, I would have been more selective in seeking a job and waiting until I got what I thought I wanted. As it turned out, however, the job with DeBevoise gave me a most liberal education in practical business and a broad background for my third and last job in business which was to come 17 years later.

Walter W.~DeBevoise was the sole owner of the business and a very sharp operator. A calm, dignified man of about 50 with graying black hair, he evoked the like of a \href{http://en.wikipedia.org/wiki/Parson}{parson} rather than a successful business man. Two relatives worked with him. Gus, a brother, was superintendent and Clarence, a nephew, his assistant. Gus was gruff and could be very tough with the workers. Clarence, on the other hand, was talkative, witty and poplular with his associates. Harry Elsebough, my immediate boss, who faced me across that long desk was a tall handsome man with a bushy brown mustache. He had a greater tendency to tell stories and play practical jokes than to get down to business. All-in-all it was a pleasant place to work despite the long ours and hard work.

The work day started at 8 am and ended at 5:30 pm. Saturday was ``half-day'' from 8 to 3:30. My duties, as outlined on that morning of September 16, 1912, were to dust the office furniture with one of those long feather dusters, open the safe and place the ledgers on the long shelf above our desk and then start posting the journal entry of the previous day's deliveries. I was also to help the girl type the billings. As it developed, however, in the day-to-day activities, I would do as Harry and the boss would do, that is pitch in on any job in the factory that required help, particularly in the shipping room tying boxes and helping load the trucks. 

The boss would arrive at the factory at about 8:15 am. He would first open the mail piled on his desk being very careful to cut the envelopes cleanly. He would pile the clean sides of the envelopes on his desk for scratch paper. Then with a few of these and a stub of a pencil he would tour the plant. When he returned he had the ``cost accounting'' data noted on the clean underside of the clipped envelopes. This data included the number of points of material in each batch of candy, the number of people on each operation, the time required to complete the jobs and all other facts gathered at first hand that would help him in estimating costs. He did not have to refer to invoices for the costs of materials. He knew just what they were. Despite this apparently inefficient method, the business prospered in those early years.

DeBevoise specialized in penny candies, the kind I used to buy in Dusenberg's store when I was a kid and were very popular because you got a lot for the money. Big pieces of chocolate-covered marshmallow, colored coconut-covered marshmallow bars, chocolate covered marshmallow pretzels, drops and the leader, Uncle Same Bars, made of ground peanuts and cheap chocolate. These bars were about the size of 5$\cent$ bars today. Then there were such novelties as Fried Eggs, little tin saucepans filled with white cream and topped with a piece of pink cream to resemble the yolk. A little tin spoon went with this ``work of art'' and the kid got it all for a penny. DeBevoise bulk candies included jelly beans, coconut bon-bons, chocolate cream drops, and the real treat of them all, hand-dipped chocolate almonds.

The candy was sold to jobbers who in turn sold to the little retail stores. There were many of these little stores. What few chain food stores in existence at that time did not sell candy and few of the chain tobacco shops did. The ice cream and confectionary stores to a large extent handled packaged and bulk candies. In retrospect, one wonders how the small candy stores catering to children met their expenses. For example, DeBevoise's Uncle Sam Bars sold to the jobber in boxes of 48 pieces for 35$\cent$. He would sell it for 38 or 40$\cent$ to the retailer who would have to make 48 separate sales for a return of 8 or 10$\cent$. A little greater return was possible for both the jobber and retailer with other candies packed 72 or 100 in boxes selling for 40 and 50$\cent$ to the jobber. In light of the narrow margin of profit in the candy business, maybe my starting salary of ten dollars a week was not so small after all.

Girls in the factory averaged about eight dollars and men about 10 dollars a week. The foreman and skilled candy makers got more, of course. Because of absences and off-season lay-offs, many of the pay envelopes -- which I helped to stuff -- contained only three or four dollars. 

Most of the girls worked in heavy sweaters in the refrigerator rooms packing chocolate candies after they were sufficiently chilled and hardened. Outside these cold rooms men lined up on each side of slow-moving endless belts and placed uncovered candies on the belts which ran through the \href{http://en.wikipedia.org/wiki/Enrober}{enrober} where they were covered with chocolate and moved on into the cold rooms.

In addition there were separate departments for the manufacture of chocolate, coconut goods, and for ``panned goods'' such as jelly beans and jawbreakers which were coated and polished in large revolving kettles. 

During the early period of my associate with DeBevoise, sanitary and working conditions were not what they are today. Food factories were far from clean, and pure food laws either non-existent or not enforced. Highly flavored and dark colored candies covered a multitude of sanitary and production sins. Working conditions were equally bad. Management was int he saddle and rode hard on labor. One example in our plant can be cited. Two men feeding an enrober quarreled and slowed down the line. In a matter of seconds, the superintendent and the foreman manhandled these workers and had them back on the line without another murmur.

Twenty years later the pendulum had swung widely in the other direction. Labor was organized and began its usurpation of power at the expense of management and the progress of industry. At this writing, labor is still in the ascendency and management greatly handicapped in efficient operations. For the future of industry and the economic progress of America, it is my hope that an equitable balance between management and labor can be reached. 

After three or four years, I began assisting in sales. I was given the South Brooklyn territory in which there were about a dozen jobbers. These were covered by trolley car and on foot in one afternoon a week. I also helped out in the Yorktown section of Manhattan on Fridays and filled in for Bill Kimberly, our city salesman, in other sections when he was ill. Another additional job, one much to my liking, was writing advertising copy for two of the trade magazines. These were but quarter page ads but gave me the opportunity to see my copy in print. Then on my 24\textsuperscript{th} birthday, April 6, 1917, the United States entered World War I.

% Timeline: 1917

Meanwhile, home in Richmond Hill our social activities were humming along as outlined in a previous part of this story. Wad Smith had taken a job with the Brooklyn Daily Eagle and we were together quite a little in Brooklyn and in Richmond Hill. I was as much at home with his friends as he was with mine. So between the two groups I was kept busy -- socially at least. As our apartment on Briggs Avenue was a bit crowded, Pop and I partitioned off two rooms in the attic with plaster board. They were fairly large rooms but bitter cold in the winter as it was not piped for heat. I would have to undress in the kitchen on these cold nights and hustle into bed under a half-ton of blankets. Later, when Grandma LaBonte came to live with us, she had the other room. She was ill and senile at times. On several occasions during the night she would fall out of bed and I would grope through to her room and lift her back into bed. Mother had a difficult time as Grandma Morris, also ill, lived with us before Grandma LaBonte came. Grandma Morris died in 1917 two days before her brother, David Swayze. Grandma LaBonte died in 1919. 

Chester had been keeping company with a little auburn haired girl Lizetta Letts for two or three years. Lizetta lived down the avenue diagonally across from our first home in Morris Park. After Chester met Lizetta, no other girl meant anything to him. They were married in 1915 in her home. It was quite a wedding. Most of our and their relatives were there. I am thinking particularly of Carrie Mather, mother's sister, her husband Raymond and their two children, Fred who was a little older than Chester and Effie, a beautiful young woman, who was about Chester's age. Fred would visit us occasionally but Effie and her father we lost track of. Uncle Raymond was a heavy drinker and years later Aunt Carrie left him and lived with the folks in Ronkonkoma as a result of that broken home. 

Perhaps because of the distance separating us, we rarely saw Mom's family. Mom had a brother, Alvie LaBonte, in Rensselaer whom I had met only once. That was when Clarence DeBevoise, Louis Pesagno, our chocolate foreman at the factory, and I spent the a weekend sightseeing in Albany. I had heard that Uncle Alvie worked with a paper manufacturing company so found his business address and dropped in for a five minute chat with him. I never met his wife or children and do not even know their names.

In 1917 Vada married Roy Rubb, a tall slim good-looking boy who lived in Richmond Hill. As Roy was not in our group of boyhood friends, I had not met him until he began courting Vada.

After the declaration of war in 1917, the boys of my age were confused. We had lived in a peaceful world most of our lives. The \href{http://en.wikipedia.org/wiki/Spanish-American_War}{Spanish-American War} in 1898 occurred when we were babies and the only other military disturbance was the expedition of the United States sent into Mexico in 1916 in an attempt to capture the Mexican Bandit \href{http://en.wikipedia.org/wiki/Pancho_Villa}{Francisco ``Pancho'' Villa}, who had raided our borders. In comparison with the war that had been going on in Europe since the summer of 1914, these were mere skirmishes. Our patriotism had been fanned to a fever heat and we wanted to offer our services voluntarily. We did not want to be drafted. We wanted to choose the branch of service we thought we could be most helpful. We certainly were confused. 

In this period of indecision, I can recall the urge to toss a coin and accept its fall on joining the army or navy. One day my business took me in the vicinity of Battery Park, Manhattan. It was a bright, warm day and I stood on the waterfront gazing at the shipping in the harbor, I made up my mind that the navy was where I belonged. 

I enlisted in the Naval Reserve as a seaman third class and was called to duty as a yeoman third class and assigned to Headquarters, Third Naval District at 280 Broadway, New York City. After all the pondering and indecision, here I was behind a desk while many of my friends were in the more romantic ends of the service -- at least, I thought then that they were romantic. Later, I realized that it was fate or good fortune that kept me out of combat, but at that age pride was a factor that influenced a feeling that we were slackers hiding behind a desk.

My duties were entirely clerical. The uniform and the discipline were the only things different from civilian life. I lived at home and commuted to New York like a civilian. I don't think I looked particularly heroic in the bell-bottom pants and gob's jacket and I was very sensitive about wearing them. Even though I was living at home, it was against regulations to don civilian clothes while there or any place. During the war the uniform was mandatory at all times. By the end of 1917, I was a yeoman first class and wore three stripes on my jacket sleeve and in May 1918 was made a Chief Yeoman and proudly wore the uniform of a chief petty officer. 

In January 1918, I learned that examinations shortly would be held for a paymaster training course at the \href{http://en.wikipedia.org/wiki/United_States_Naval_Academy}{Naval Academy in Annapolis, Maryland} starting on August 1. This seemed to offer a real opportunity to get away from New York and perhaps get a commission and an assignment to sea duty. With the recommendation of my commanding officer and letters from DeBevoise, United States Title and the principal of Commercial High, I applied for permission to take the examination. Permission was granted and I took and passed the examination.

Much to my disappointment, I was not included in the Annapolis group. I did so much want the opportunity to take training in that historic academy. But my disappointment was tempered with the announcement a short time later that another training course, United States Navy Officer Material School for the Pay Corps, was to be established at Princeton University and that I would be included. Now I was happy. I was about the remove what I thought was the stigma of desk duty in the City and, at the same time had proven to myself that I could compete successfully in an effort to improve my status. It was the first real test of competitive competence outside of school and sports. My folks were proud. And I hoped a certain charming young lady I had met a short time before shared that pride. 

\chapter{A Certain Charming Young Lady}

In the spring on 1918, Art Wenige and I dropped in at a dance for servicemen (Art was an Army trainee at \href{http://en.wikipedia.org/wiki/Camp_Upton}{Camp Upton}) given by the USO in Brooklyn. There Art met a young lady in a Red Cross Nurse's univorm. When Art suggested a date, she agreed on the condition that a friend of hers come along if I would care to make it a double date. I agreed but wondered who was being imposed on me as I had had some experience on double dates and never seemed to pick a winner. 

We kept the date. I can't remember where we went or what we did because I was in a daze. I had met THE girl -- Mabel Norseen, an 18 year old slim blond girl with a ``peaches and cream'' complexion. She lived at 563 10\textsuperscript{th} Street in the Park Slope section of Brooklyn and I can remember leaving her that night after she promised me another date and somehow or other finding my way home to Richmond Hill. I was really ``hit hard.''

I wondered what she thought about me, whether she felt that I was the ``boy of her dreams.'' I doubt that she did because I was far from a romantic figure in that gob's uniform which accentuated my thin, pale face and my slightly bowed legs. These are the thoughts that bothered me after that first meeting. But she had agreed to another date so maybe she did like me. Shortly after that eventful first meeting, I proudly called on her wearing my chief petty officer's uniform which gave me a lot more confidence in my burning desire to make a favorable impression. Maybe I did because we were together several times before I left for the paymaster's school at Princeton with her promise to write after. 

Along with 100 or more hopefuls (perhaps 200), I reported to the Naval officer in charge of training at Princeton early in September 1918. We were billeted in beautiful Graduate College at the far end of the Princeton campus and I was thrilled with the place. At long last I was at ``college'' or at least in the environment of a college. I was billeted with three other ``M's'' -- Murdock, Moses and Morrisey -- in a room in the graduate building. It was an oddly assorted group. Murdock was a Harvard man, later to become a professor there; Morrisey, a tough little Irishman from New York City; and Moses, a quiet spoken boy of Jewish descent and also from the City. 

The training was not too difficult. It consisted of lectures on navigation and on subjects concerned with naval regulations and paymaster's duties with on period a day devoted to actual military training on the parade grounds. I cannot understand why we took such military training when our jobs were to be on paper work. It must have been for the discipline factors. 













\end{document}